---
title: "Reproducible Research: Peer Assessment 1"
date: "`r Sys.Date()`"
output: html_document
keep_md: true
---
## Documenting the current environment

I know this was not part of the template for this assignment, but Prof Peng 
mentions in the week 3 lectures that part of the Reproducible Research workflow 
is to document the environment that the analysis is run in (just in case there 
are issues that relate to a specific version).  

I have therefore put in this 
section to remind myself later how to do this...

```{r session_info}
# first load the required libraries
library(dplyr)

# then generate the snapshot of the environment used
sessionInfo()
```

## Loading and preprocessing the data

Need to unzip the supplied file, read it into memory, and then clean up the
unzipped files.

```{r read_data}
unzip("activity.zip", exdir = "unzipped")
activity <- read.csv("unzipped/activity.csv")
file.remove(c(list.files(path = "unzipped", recursive = TRUE, full.names = TRUE),
                      "unzipped"))
```

## What is mean total number of steps taken per day?

There are a lot of NA's in the step count:
```{r display_na}
str(activity)
```
I will therefore remove the NA's before calculating the daily total step count:
```{r daily_totals}
daily_totals <- activity %>% group_by(date) %>% na.omit %>% summarise(total = sum(steps))
daily_totals
```
This allows me to plot the required histogram:
```{r daily_totals_hist}
hist(daily_totals$total, breaks=25L, col="dark blue", 
     main="Frequency distribution of the total number of steps per day")
```

The mean number of steps per day is `r format(mean(daily_totals$total), scientific=F, big.mark=",")` steps.  

The median number of steps per day is `r format(median(daily_totals$total), scientific=F, big.mark=",")` steps.  

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
